STAT 672 Final Project:\\ Stochastic Gradient Descent

### 2018-04-18 paper/paper.tex
minor changes

Consider a feature matrix $\bm{X}$ and outcome variable $\bm{Y}$, and the task
of fitting coefficients in a model to predict $Y_i$ based on $\bm{X}_i$. 
Assuming that particular model elements follow a
particular distribution aids the estimation of coefficients. 
For example, assumptions in ordinary least squares (OLS)
regression---assumptions that readers almost certainly are familiar with and so will not be
repeated here---allow a closed form solution, $\hat{\bm{\beta}} = (\bm{X}'\bm{X})^{-1}\bm{X}'\bm{Y}$.

@@

$(t)$ refers to a particular iteration, $\bm{J}$ is the Jacobian,
and $\bm{H}$ is the Hessian. The practicality of Newton-Raphson thus depends on whether it is convenient to
find $\bm{J}$ and in particular $\bm{H}$. It is convenient for logistic regression
because parametric and independent-and-identically-distributed (IID) assumptions
mean $l$ is a simple sum of the log probability distribution
function for each observation. We ``know'' (assume) the form of this PDF and so are
confident that the second derivative exists and is not onerous.
In non-parametric settings, we often cannot be so certain and face the possibility of
$\bm{H}$ being non-existent or cumbersome.

@@

We could apply the GD algorithm above to find the optimal $w$.

@@

The above proof is for GD. A rigorous proof is not given for why SGD also
converges to the optimal value. Informally, we can note that the above proof
says given infinite $t$ and appropriate $\gamma$, the iterative algorithm will
always converge to the optimal value. SGD is doing the same thing as GD, 
just with a single observation per iteration 
rather than the entire dataset per iteration. Since SGD is using less information
per iteration, we would expect the convergence to require more iterations. But
per (9) we expect it to \textit{eventually} arrive at the optimal solution.
If so, the difference between GD and SGD must then primarily revolve
around convergence speed, not the final value that is converged to.

>>>>

Consider a feature matrix $\bm{X}$ and outcome variable $\bm{Y}$, and the task
of fitting coefficients in a model to predict $Y_i$ based on $\bm{X}_i$. 
Assuming that particular model elements follow a
known statistical distribution aids the estimation of coefficients. 
For example, assumptions in ordinary least squares (OLS)
regression---assumptions that readers almost certainly are familiar with and so will not be
repeated here---allow a closed form solution, $\hat{\bm{\beta}} = (\bm{X}'\bm{X})^{-1}\bm{X}'\bm{Y}$.

@@

$(t)$ refers to a particular iteration, $\bm{J}$ is the Jacobian of the
log likelihood function $l$,
and $\bm{H}$ is the Hessian of $l$. The practicality of Newton-Raphson thus depends on whether it is convenient to
find $\bm{J}$ and in particular $\bm{H}$. It is convenient for logistic regression
because parametric and independent-and-identically-distributed (IID) assumptions
mean $l$ is a simple sum of the log probability distribution
function (PDF) for each observation. We ``know'' (assume) the form of this PDF and so are
confident that the second derivative exists and is not onerous.
In non-parametric settings, we often cannot be so certain and face the possibility of
$\bm{H}$ being non-existent or cumbersome.

@@

We could apply the GD algorithm above to find $\hat{w}$.

@@

The above proof is for GD. Proof is not given for why SGD also
converges to the optimal value. Informally, we can note that the above proof
says given infinite $t$ and appropriate $\gamma$, the iterative algorithm will
always converge to the optimal value. SGD is doing the same thing as GD, 
just with a single observation per iteration 
rather than the entire dataset per iteration. Since SGD is using less information
per iteration, we would expect the convergence to require more iterations. But
per (9) we expect it to \textit{eventually} arrive at the optimal solution.
If so, the difference between GD and SGD must then primarily revolve
around convergence speed, not the final value that is converged to.
